---
title: "Text Visualization"
description: |
  We will explore ways to visualize text data using R in this blog.
author:
  - name: Alicia Zhang
    url: https://www.linkedin.com/in/alicia-zhang-22a1a6140/
    affiliation: School of Computing and Information Systems, Singapore Management University
    affiliation_url: https://scis.smu.edu.sg/
date: 07-11-2021
output:
  distill::distill_article:
    toc: true
    toc_depth: 3
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.retina=3,
                      echo = TRUE,
                      eval = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

# Installing and launching R packages 

```{r}

packages = c("tidytext","widyr","wordcloud",
             "DT","ggwordcloud","textplot",
             "lubridate","hms","tidyverse",
             "tidygraph","ggraph","igraph")
for (p in packages){
  if(!require(p,character.only = T)){
    install.packages(p)
  }
  library(p,character.only = T)
}
```

# Import Multiple Text Files from Multiple Folders 
**Step 1: Creating a folder list**

```{r}
news20 <- "data/20news/"
```

**Step 2: Define a function to read all files from a folder into a data frame**

```{r}
read_folder <- function(infolder){
  tibble(file = dir(infolder,
                    full.names=TRUE)) %>%
    mutate(text=map(file,
                    read_lines)) %>%
    transmute(id=basename(file),
              text)%>%
    unnest(text)
}
```

**Step 3: Reading in all the messages from the 20news folder**

```{r}
raw_text <- tibble(folder=dir(news20,
                              full.names = TRUE)) %>%
  mutate(folder_out = map(folder,
                          read_folder))%>%
  unnest(cols = c(folder_out))%>%
  transmute(newsgroup=basename(folder),
            id,text)
write_rds(raw_text,"data/rds/news20.rds")
```

# Initial EDA

```{r}
raw_text %>%
  group_by(newsgroup) %>%
  summarise(messages = n_distinct(id))%>%
  ggplot(aes(messages,newsgroup))+
  geom_col(fill="lightblue")+
  labs(y=NULL)

```

# Cleaning Text Data

### Removing header and automated email signitures 

```{r}
cleaned_text <- raw_text %>%
  group_by(newsgroup,id) %>%
  filter(cumsum(text == "")>0,
         cumsum(str_detect(
           text,"^--")) == 0) %>%
  ungroup()
```


### Removing lines with nested text representing quotes from other users.

```{r}
cleaned_text <- cleaned_text%>%
  filter(str_detect(text,"^[^>]+[A-Za-z\\d]")
         |text == "",
         !str_detect(text,
                     "writes(:|\\.\\.\\.)$"),
         !str_detect(text,
                     "^In article <"))

```

# Text Data Processing

```{r}
usenet_words <- cleaned_text %>%
unnest_tokens(word, text) %>%
filter(str_detect(word, "[a-z']$"),
!word %in% stop_words$word)
```

### check the frequency of words

```{r}
usenet_words %>%
  count(word,sort=TRUE)

```

### count words within by newsgroup

```{r}
words_by_newsgroup <- usenet_words %>%
  count(newsgroup, word, sort = TRUE) %>%
  ungroup()

```


#  Visualising Words in newsgroups

### Using wordcloud package

```{r}
wordcloud(words_by_newsgroup$word,
          words_by_newsgroup$n,
          max.words = 300)
```

## Computing tf-idf within newsgroups

```{r}
tf_idf <- words_by_newsgroup %>%
  bind_tf_idf(word, newsgroup, n) %>%
  arrange(desc(tf_idf))
```

## Visualising tf-idf as interactive table

```{r}
DT::datatable(tf_idf,filter="top")%>%
  formatRound(columns = c('tf','idf','tf_idf'),
              digits=3)%>%
  formatStyle(0,target='row',lineHeight='25%')
``` 

## Visualising tf-idf within newsgroups

```{r}
tf_idf%>%
  filter(str_detect(newsgroup,"^sci\\."))%>%
  group_by(newsgroup)%>%
  slice_max(tf_idf,
            n=12)%>%
  ungroup()%>%
  mutate(word=reorder(word,tf_idf))%>%
  ggplot(aes(tf_idf,
             word,
             fill=newsgroup))+
  geom_col(show.legend=FALSE)+
  facet_wrap(~newsgroup,
             scales="free")+
  labs(x="tf-idf",
       y=NULL)

``` 


## Counting and correlating pairs of words with the widyr 

```{r}
newsgroup_cors <- words_by_newsgroup%>%
  pairwise_cor(newsgroup,
               word,
               n,
               sort=TRUE)

``` 

## Visualising correlation as a network

```{r}
set.seed(2017)

newsgroup_cors%>%
  filter(correlation > .025)%>%
  graph_from_data_frame()%>%
  ggraph(layout="fr") +
  geom_edge_link(aes(alpha=correlation,
                     width=correlation)) +
  geom_node_point(size=6,
                  color="lightblue")+
  geom_node_text(aes(label=name),
                 color = "red",
                 repel = TRUE)+
  theme_void()

``` 

# Bigram

```{r}
bigrams <- cleaned_text%>%
  unnest_tokens(bigram,
                text,
                token = "ngrams",
                n=2)

``` 


# Counting bigrams

```{r}
bigrams_count <- bigrams %>%
  filter(bigram != 'NA') %>%
  count(bigram, sort = TRUE)

``` 

## Cleaning bigram

```{r}
bigrams_separated <- bigrams %>%
  filter(bigram != 'NA') %>%
  separate(bigram, c("word1", "word2"),
           sep = " ")

bigrams_filtered <- bigrams_separated %>%
  filter(!word1 %in% stop_words$word) %>%
  filter(!word2 %in% stop_words$word)
```


## counting bigram

```{r}
bigram_counts <- bigrams_filtered %>%
  count(word1, word2, sort = TRUE)
``` 


## create a network graph from bigrams 

```{r}
bigram_graph <- bigram_counts %>%
  filter(n > 3) %>%
  graph_from_data_frame()
bigram_graph
``` 


## create a network graph from bigrams 

```{r}
set.seed(1234)
ggraph(bigram_graph, layout = "fr") +
  geom_edge_link() +
  geom_node_point() +
  geom_node_text(aes(label = name),
                 vjust = 1,
                 hjust = 1)
``` 


## improved version

```{r}
set.seed(1234)
a <- grid::arrow(type = "closed",
                 length = unit(.15,
                               "inches"))
ggraph(bigram_graph,
       layout = "fr") +
  geom_edge_link(aes(edge_alpha = n),
                 show.legend = FALSE,
                 arrow = a,
                 end_cap = circle(.07,
                                  'inches')) +
  geom_node_point(color = "lightblue",
                  size = 5) +
  geom_node_text(aes(label = name),
                 vjust = 1,
                 hjust = 1) +
  theme_void()
```


