---
title: "Network Data Visualisation and Analysis"
description: |
  We will explore ways to visualize network data using R in this post.
author:
  - name: Alicia Zhang
    url: https://www.linkedin.com/in/alicia-zhang-22a1a6140/
    affiliation: School of Computing and Information Systems, Singapore Management University
    affiliation_url: https://scis.smu.edu.sg/
date: 07-10-2021
output:
  distill::distill_article:
    toc: true
    toc_depth: 3
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, error = FALSE)
```

# R graph objects
    + network: Classes for Relational Data.
    + igraph: Network Analysis and Visualzation. It's a wrapper of tidygraph.
    + tidygraph: A Tidy API for Graph Manipulation.
    + ggraph and visNetwork: to visualize graphs. 
    
# Installing and loading the required libraries

Use the below code to install and load packages. 

```{r}

packages = c('igraph','tidygraph',
             'ggraph','visNetwork',
             'lubridate','clock',
             'tidyverse')
for (p in packages){
  if(!require(p,character.only = T)){
    install.packages(p)
  }
  library(p,character.only = T)
}
```

# Importing network data from files 

```{r}
GAStech_nodes <- read_csv("data/GAStech_email_node.csv")
GAStech_edges <- read_csv("data/GAStech_email_edge-v2.csv")

```

## Reviewing the imported data 

SentDate is a character data instead of date data.
```{r}
glimpse(GAStech_edges)

```

# Wrangling time
dmy() transforms the SentDate to Date data type
wday() returns the day of the week as a decimal number or an ordered factor if label is TRUE

```{r}
GAStech_edges$SentDate = dmy(GAStech_edges$SentDate)
GAStech_edges$Weekday = wday(GAStech_edges$SentDate,
                             label = TRUE,
                             abbr = FALSE)

```

# Wrangling attributes

```{r}
GAStech_edges_aggregated <- GAStech_edges %>%
  filter(MainSubject == "Work related") %>%
  group_by(source,target, Weekday)%>%
  summarize(Weight = n())%>%
  filter(source!=target)%>% #exclude people who wrote to themselves
  filter(Weight > 1)%>% # exclude all nodes with weights less or equal to 1
  ungroup
```

## Reviewing the new data set
```{r}
glimpse(GAStech_edges_aggregated)
```

# Using tbl_graph() to build tidygraph data model


```{r}
GAStech_graph <- tbl_graph(nodes = GAStech_nodes,
                           edges = GAStech_edges_aggregated,
                           directed = TRUE)
GAStech_graph
```

## Reviewing the output tidygraph's graph object

```{r}
GAStech_graph
```

# Plotting network graphs

## Plotting a basic network graph

```{r}
ggraph(GAStech_graph) +
  geom_edge_link() +
  geom_node_point()
```

## Changing the default newwork graph theme

```{r}
g <- ggraph(GAStech_graph) +
  geom_edge_link(aes()) +
  geom_node_point(aes()) 
g + theme_graph() # remove grey background color
```
## Changing the coloring of the plot

```{r}
g <- ggraph(GAStech_graph) +
  geom_edge_link(aes(colour=("grey50"))) +
  geom_node_point(aes(colour="grey40")) 
g + theme_graph()
```

## Fruchterman and Reingold layout

```{r}
g <- ggraph(GAStech_graph,
            layout = "fr") +
  geom_edge_link(aes()) +
  geom_node_point(aes()) 
g + theme_graph()
```


## Modifying network nodes

```{r}
g <- ggraph(GAStech_graph,
            layout = "nicely") +
  geom_edge_link(aes()) +
  geom_node_point(aes(colour = Department,
                      size=3)) 
g + theme_graph()
```


## Modifying network edges

```{r}
g <- ggraph(GAStech_graph,
            layout = "nicely") +
  geom_edge_link(aes(width=Weight),
                 alpha=0.2) +
  scale_edge_width(range=c(0.1,5))+
  geom_node_point(aes(colour=Department),
                  size =3)
g + theme_graph()
```

## Working with facet_edges()

```{r}
set_graph_style()

g <- ggraph(GAStech_graph,
            layout = "nicely") +
  geom_edge_link(aes(width=Weight),
                 alpha=0.2) +
  scale_edge_width(range=c(0.1,5))+
  geom_node_point(aes(colour=Department),
                  size =2)
g + facet_edges(~Weekday)
```


## change legend position 

```{r}
set_graph_style()

g <- ggraph(GAStech_graph,
            layout = "nicely") +
  geom_edge_link(aes(width=Weight),
                 alpha=0.2) +
  scale_edge_width(range=c(0.1,5))+
  geom_node_point(aes(colour=Department),
                  size =2)+
  theme(legend.position = "bottom")

g + facet_edges(~Weekday)

```

## A framed facet graph
```{r}
set_graph_style()

g <- ggraph(GAStech_graph,
            layout = "nicely") +
  geom_edge_link(aes(width=Weight),
                 alpha=0.2) +
  scale_edge_width(range=c(0.1,5))+
  geom_node_point(aes(colour=Department),
                  size =2)

g + facet_edges(~Weekday)+
  th_foreground(foreground = "grey80",
                border = TRUE)+
  theme(legend.position="bottom")

```


## Working with facet_nodes()
```{r}
set_graph_style()

g <- ggraph(GAStech_graph,
            layout = "nicely") +
  geom_edge_link(aes(width=Weight),
                 alpha=0.2) +
  scale_edge_width(range=c(0.1,5))+
  geom_node_point(aes(colour=Department),
                  size =2)

g + facet_nodes(~Department)+
  th_foreground(foreground = "grey80",
                border = TRUE)+
  theme(legend.position="bottom")

```

# Network Metrics Analysis
##  Computing centrality indices
```{r}
g <- GAStech_graph %>%
  mutate(betweenness_centrality = centrality_betweenness())%>%
  ggraph(layout="fr")+
  geom_edge_link(aes(width=Weight),
                 alpha=0.2) +
  scale_edge_width(range=c(0.1,5))+
  geom_node_point(aes(colour=Department,
                  size = betweenness_centrality))

g + theme_graph()

```

## Visualising network metrics
```{r}
g <- GAStech_graph %>%
  ggraph(layout = "fr") +
 geom_edge_link(aes(width=Weight),
                 alpha=0.2) +
  scale_edge_width(range=c(0.1,5))+
  geom_node_point(aes(colour=Department,
                      size=centrality_betweenness()))

g + theme_graph()

```

## Visualising Community
```{r}
g <- GAStech_graph %>%
  mutate(community = as.factor(group_edge_betweenness(weights=Weight, directed = TRUE)))%>%
  ggraph(layout = "fr") +
 geom_edge_link(aes(width=Weight),
                 alpha=0.2) +
  scale_edge_width(range=c(0.1,5))+
  geom_node_point(aes(colour=community))

g + theme_graph()
```

# Building Interactive Network Graph with VisNetwork
## Data preparation
```{r}

GAStech_edges_aggregated <- GAStech_edges %>%
  left_join(GAStech_nodes, by = c("sourceLabel"="label"))%>%
  rename(from=id)%>%
  left_join(GAStech_nodes, by = c("targetLabel"="label"))%>%
  rename(to=id)%>%
  filter(MainSubject=="Work related")%>%
  group_by(from,to)%>%
  summarise(weight=n())%>%
  filter(from!=to)%>%
  filter(weight>1)%>%
  ungroup()
```

## Plotting the first interactive network graph
```{r}
visNetwork(GAStech_nodes,
           GAStech_edges_aggregated)%>%
  visIgraphLayout("layout_with_fr") # can choose from igraph layouts
```

## Working with visual attributes - Nodes
```{r}
GAStech_nodes <- GAStech_nodes %>%
  rename(group=Department)
```

```{r}
visNetwork(GAStech_nodes,
           GAStech_edges_aggregated)%>%
  visIgraphLayout("layout_with_fr")%>%
  visLegend()%>%
  visLayout(randomSeed=123) #set random
```

## Interactivity
```{r}
visNetwork(GAStech_nodes,
           GAStech_edges_aggregated)%>%
  visIgraphLayout("layout_with_fr")%>%
  visOptions(highlightNearest = TRUE,
             nodesIdSelection = TRUE)%>%
  visLegend()%>%
  visLayout(randomSeed=123) 

```